Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-12-14 paper/main.tex
Removed \ac from sections

In light of the ongoing NIST competition for selecting the new, standard
cryptographic hash function named SHA-3, we tried out the 14 second
round candidates in the Tahoe-LAFS file system using automated tests to measure
the performance differences. Only the immutable file support was tested.

Since the Tahoe-LAFS is programmed in Python, we had to make bindings to the
SHA-3 implementations written in C, before making the necessary changes to the
Tahoe-LAFS code. Several tasks where automated using scripts, to avoid doing
manual work of distributing changes to the test grid.

In general, four of the candidates perform better than the currently used
SHA-256 -- BLAKE, BlueMidnightWish, Shabal and Simd.
By utilizing the performance improvements given by SHA-3, Tahoe-LAFS could
save $5\%$ on upload operations by changing hashing function to the best
performing candidate.

@@

\subsubsection{\ac{NIST} evaluation criteria for \ac{SHA}-3}

@@

\subsubsection{\ac{NIST} \ac{SHA}-3 \ac{API}} \ac{NIST} required that every
submission should conform to a specified \ac{API}\cite{s_fedreg}. The \ac{API}
specification \cite{s_nistapi} states in shortness that it requires every
candidate to implement four functions:

@@

\subsection{\ac{SUPERCOP}}
\ac{SUPERCOP} is a toolkit developed by VAMPIRE lab for measuring performance
of cryptographic software \cite{s_supercop}. In relation to hash functions,
\ac{SUPERCOP} measures the following:

@@

\subsubsection{\ac{SUPERCOP} \ac{API}}
\ac{SUPERCOP} specifies an \ac{API} which all submitted implementations must
conform to \cite{s_supercopapi}. This \ac{API} specifies the naming and
organization of files, and that a submission must include a function
{\verb crypto_hash() } which is similar to the function {\verb Hash() } in the
\ac{NIST} \ac{SHA}-3 \ac{API}. The only difference being the naming of
data types and that the {\verb crypto_hash() } function does not need an input
of how long the output of the hash function should be. Instead double or
different submissions are used for different output lengths.

@@

\subsubsection{\ac{Distutils}}
We use \ac{Distutils}\footnote{\url{http://docs.python.org/distutils/}} to
build and install our code. The alternative to doing this would be to manually
transform Cython files to C-files, build the shared library from this and other
relevant C-files and then install by copying into the correct system directory
where Python libraries should be stored. \ac{Distutils} also support expansion
through self defined commands, something we have utilised to do Unit Testing.
Options and configuration for \ac{Distutils} is done through a file
{\verb setup.py}. Relevant configuration for compiling the echo extension, and
installing SHA3lib can bee seen in Listing \ref{lst:setup_py}.

@@

\section{Modifications to the Tahoe-\ac{LAFS} Code}
\label{sec:modtahoe}

@@

\section{Measurements of Tahoe data}
\label{sec:results:tahoedata}

@@

\section{Measurements of Tahoe operations}
The sum of the amount of time it took for the underlying hash library to
calculate the hash operations are exhibited in Figures \ref{fig:graph:1001b},
\ref{fig:graph:1001kb}, \ref{fig:graph:501mb}, \ref{fig:graph:5100mb} and
\ref{fig:graph:11gb}. We call this \emph{Time spent hashing}.

@@

\section{Validity of \ac{SUPERCOP} results}

@@

\subsection{Impact of double \ac{SHA}-256}
The impact of using double \ac{SHA}-256 compared to single is surprisingly
small, even if you use the best-case estimate from Table
\ref{tbl:supercop:sha256}. For 1B, 1KB and 1MB files the data hashed is almost
exclusively 106B or less. If we assume data is hashed at the performance of 576
bytes, single \ac{SHA}-256 should outperform double \ac{SHA}-256 by about 10\%.
For 1B files this difference is eaten up by external factors (other
processes), the same goes for 1KB files. For 1MB files, there seems to be a
difference, however this difference is so small that we deem it unimportant.
For the larger files the performance penalty is even smaller, so all in all we
deem the performance penalty for Tahoe-\ac{LAFS} for using double \ac{SHA}-256
as opposed to single unimportant.

>>>>

In light of the ongoing NIST competition for selecting the new, standard
cryptographic hash function named SHA-3, 14 of the second round candidates were
tried out in the Tahoe-LAFS file system using automated tests to measure the
performance differences. Only the immutable file support was tested.

Since the Tahoe-LAFS is programmed in Python, bindings to the SHA-3
implementations written in C had to be made , before making the necessary
changes to the Tahoe-LAFS code. Several tasks where automated using scripts, to
avoid doing manual work of distributing changes to the test grid.

In general, four of the candidates perform better than the currently used
SHA-256 -- BLAKE, BlueMidnightWish, Shabal and Simd.  By utilizing the
performance improvements given by SHA-3, Tahoe-LAFS could currently save $5\%$
on upload operations by changing hashing function to the best performing
candidate.

@@

\subsubsection{NIST evaluation criteria for SHA-3}

@@

\subsubsection{NIST SHA-3 API}

\ac{NIST} required that every
submission should conform to a specified \ac{API}\cite{s_fedreg}. The \ac{API}
specification \cite{s_nistapi} states in shortness that it requires every
candidate to implement four functions:

@@

\subsection{SUPERCOP}
\ac{SUPERCOP} is a toolkit developed by VAMPIRE lab for measuring performance
of cryptographic software \cite{s_supercop}. In relation to hash functions,
\ac{SUPERCOP} measures the following:

@@

\subsubsection{SUPERCOP API}
\ac{SUPERCOP} specifies an \ac{API} which all submitted implementations must
conform to \cite{s_supercopapi}. This \ac{API} specifies the naming and
organization of files, and that a submission must include a function
{\verb crypto_hash() } which is similar to the function {\verb Hash() } in the
\ac{NIST} \ac{SHA}-3 \ac{API}. The only difference being the naming of
data types and that the {\verb crypto_hash() } function does not need an input
of how long the output of the hash function should be. Instead double or
different submissions are used for different output lengths.

@@

\subsubsection{Distutils}
We use \ac{Distutils}\footnote{\url{http://docs.python.org/distutils/}} to
build and install our code. The alternative to doing this would be to manually
transform Cython files to C-files, build the shared library from this and other
relevant C-files and then install by copying into the correct system directory
where Python libraries should be stored. \ac{Distutils} also support expansion
through self defined commands, something we have utilised to do Unit Testing.
Options and configuration for \ac{Distutils} is done through a file
{\verb setup.py}. Relevant configuration for compiling the echo extension, and
installing SHA3lib can bee seen in Listing \ref{lst:setup_py}.

@@

\section{Modifications to the Tahoe-LAFS Code}
\label{sec:modtahoe}

@@

\section{Measurements of internal Tahoe hashing sizes}
\label{sec:results:tahoedata}

@@

\section{Measurements of Tahoe operations Using Test Vectors}
The sum of the amount of time it took for the underlying hash library to
calculate the hash operations are exhibited in Figures \ref{fig:graph:1001b},
\ref{fig:graph:1001kb}, \ref{fig:graph:501mb}, \ref{fig:graph:5100mb} and
\ref{fig:graph:11gb}. We call this \emph{Time spent hashing}.

@@

\section{Validity of SUPERCOP results}

@@

\subsection{Impact of double SHA-256}
The impact of using double \ac{SHA}-256 compared to single is surprisingly
small, even if you use the best-case estimate from Table
\ref{tbl:supercop:sha256}. For 1B, 1KB and 1MB files the data hashed is almost
exclusively 106B or less. If we assume data is hashed at the performance of 576
bytes, single \ac{SHA}-256 should outperform double \ac{SHA}-256 by about 10\%.
For 1B files this difference is eaten up by external factors (other
processes), the same goes for 1KB files. For 1MB files, there seems to be a
difference, however this difference is so small that we deem it unimportant.
For the larger files the performance penalty is even smaller, so all in all we
deem the performance penalty for Tahoe-\ac{LAFS} for using double \ac{SHA}-256
as opposed to single unimportant.
